%!TEX root = io2d.tex
\rSec0 [\iotwod.brush] {Class \tcode{brush}}

\rSec1 [\iotwod.brush.synopsis] {\tcode{brush} synopsis}

\begin{codeblock}
namespace std { namespace experimental { namespace io2d { inline namespace v1 {
  class brush {
  public:
    // construct/copy/move/destroy:
    brush() = delete;
    brush(const brush&) noexcept;
    brush& operator=(const brush&) noexcept;
    brush(brush&& other) noexcept;
    brush& operator=(brush&& other) noexcept;

    // \ref{\iotwod.brush.modifiers}, modifiers:
    void extend(::std::experimental::io2d::extend e) noexcept;
    void filter(::std::experimental::io2d::filter f) noexcept;
    void matrix(const matrix_2d& m) noexcept;

    // \ref{\iotwod.brush.observers}, observers:
    ::std::experimental::io2d::extend extend() const noexcept;
    ::std::experimental::io2d::filter filter() const noexcept;
    matrix_2d matrix() const noexcept;
    brush_type type() const noexcept;

  private:
    ::std::experimental::io2d::extend _Extend; // \expos
    ::std::experimental::io2d::filter _Filter; // \expos
    matrix_2d _Matrix;                         // \expos
    brush_type _Brush_type;                    // \expos
  };
} } } }
\end{codeblock}

\rSec1 [\iotwod.brush.intro] {\tcode{brush} Description}

\pnum
\indexlibrary{\idxcode{brush}}
The class \tcode{brush} describes an opaque resource wrapper for color and alpha data.

\pnum
A \tcode{brush} object's color and alpha data is immutable.

\pnum
A \tcode{brush} object also has mutable \tcode{extend}, \tcode{filter}, and \tcode{matrix_2d} data which affects how the color and alpha data is used and interpreted.

\pnum
Lastly, a \tcode{brush} object has immutable \tcode{brush_type} data which indicates the type of brush factory which was used in creating the \tcode{brush} object.

\pnum
A \tcode{brush} object is usable as the color and alpha source for a surface's stroke, fill, mask, paint, and text rendering operations. See \ref{\iotwod.surface.modifiers.render} for more information.

\pnum
It is also usable as the mask for a surface's mask operation. Its alpha channel is used as the mask such that only brush types which have a non-uniform alpha channel are suitable for use as a mask. See the description of \tcode{surface::mask} (\ref{\iotwod.surface.modifiers.render}) for more information.

\rSec1 [\iotwod.brush.modifiers]{\tcode{brush} modifiers}

\indexlibrary{\idxcode{brush}!\idxcode{extend}}
\indexlibrary{\idxcode{extend}!\idxcode{brush}}
\begin{itemdecl}
    void extend(::std::experimental::io2d::extend e) noexcept;
\end{itemdecl}
\begin{itemdescr}
	\pnum
	\postconditions
	\tcode{_Extend == e}.
	
\end{itemdescr}

\indexlibrary{\idxcode{brush}!\idxcode{filter}}
\indexlibrary{\idxcode{filter}!\idxcode{brush}}
\begin{itemdecl}
    void filter(::std::experimental::io2d::filter f) noexcept;
\end{itemdecl}
\begin{itemdescr}
	\pnum
	\postconditions
	\tcode{_Filter == f}.
	
\end{itemdescr}

\indexlibrary{\idxcode{brush}!\idxcode{matrix}}
\indexlibrary{\idxcode{matrix}!\idxcode{brush}}
\begin{itemdecl}
    void matrix(const matrix_2d& m) noexcept;
\end{itemdecl}
\begin{itemdescr}
	\pnum
	\postconditions
	\tcode{_Matrix == m}.
	
\end{itemdescr}

\rSec1 [\iotwod.brush.observers]{\tcode{brush} observers}

\indexlibrary{\idxcode{brush}!\idxcode{}}
\indexlibrary{\idxcode{}!\idxcode{brush}}
\begin{itemdecl}
    ::std::experimental::io2d::extend extend() const noexcept;
\end{itemdecl}
\begin{itemdescr}
	\pnum
	\returns
	\tcode{_Extend}.

\end{itemdescr}

\indexlibrary{\idxcode{brush}!\idxcode{}}
\indexlibrary{\idxcode{}!\idxcode{brush}}
\begin{itemdecl}
    ::std::experimental::io2d::filter filter() const noexcept;
\end{itemdecl}
\begin{itemdescr}
	\pnum
	\returns
	\tcode{_Filter}.

\end{itemdescr}

\indexlibrary{\idxcode{brush}!\idxcode{}}
\indexlibrary{\idxcode{}!\idxcode{brush}}
\begin{itemdecl}
    matrix_2d matrix() const noexcept;
\end{itemdecl}
\begin{itemdescr}
	\pnum
	\returns
	\tcode{_Matrix}.

\end{itemdescr}

\indexlibrary{\idxcode{brush}!\idxcode{type}}
\indexlibrary{\idxcode{type}!\idxcode{brush}}
\begin{itemdecl}
    brush_type type() const noexcept;
\end{itemdecl}
\begin{itemdescr}
	\pnum
	\returns
	\tcode{_Brush_type}.

\end{itemdescr}
