%!TEX root = io2d.tex
\rSec0 [\iotwod.io2d.error] {Enum class \tcode{io2d_error}}

\rSec1 [\iotwod.io2d.error.summary] {\tcode{io2d_error} Summary}
\pnum
The \tcode{io2d_error} enum class is used as an \tcode{error_condition} 
associated with \tcode{io2d_error_category}. See 
Table~\ref{tab:\iotwod.io2d.error.meanings} for the meaning of each
status code.

\rSec1 [\iotwod.io2d.error.synopsis] {\tcode{io2d_error} Synopsis}

\indexlibrary{\idxcode{io2d_error}}%
\begin{codeblock}
namespace std { namespace experimental { namespace io2d { inline namespace v1 {
  enum class io2d_error {
    success,
    no_memory,
    invalid_restore,
    invalid_pop_group,
    no_current_point,
    invalid_matrix,
    invalid_status,
    null_pointer,
    invalid_string,
    invalid_path_data,
    read_error,
    write_error,
    surface_finished,
    surface_type_mismatch,
    pattern_type_mismatch,
    invalid_content,
    invalid_format,
    invalid_visual,
    file_not_found,
    invalid_dash,
    invalid_dsc_comment,
    invalid_index,
    clip_not_representable,
    temp_file_error,
    invalid_stride,
    font_type_mismatch,
    user_font_immutable,
    user_font_error,
    negative_count,
    invalid_clusters,
    invalid_slant,
    invalid_weight,
    invalid_size,
    user_font_not_implemented,
    device_type_mismatch,
    device_error,
    invalid_mesh_construction,
    device_finished,
    last_status
  };
} } } } // namespaces std::experimental::io2d::v1
\end{codeblock}

\rSec1 [\iotwod.io2d.error.enumerators] {\tcode{io2d_error} Enumerators}

\begin{libreqtab2}
  {\tcode{io2d_error} enumerator meanings}
  {tab:\iotwod.io2d.error.meanings}
  \\ \topline
  \lhdr{Enumerator}
  & \rhdr{Meaning}
  \\ \capsep
  \endfirsthead
  \continuedcaption\\
  \hline
  \lhdr{Enumerator}
  & \rhdr{Meaning}
  \\ \capsep
  \endhead
 \tcode{success}
 & The operation completed successfully.
 \\
 \tcode{no_memory}
 & The operation failed due to insufficient memory.
 \enternote
 Deprecated. Implementations should throw an exception of type
 \tcode{std::bad_alloc} instead.
 \exitnote
 \\
 \tcode{invalid_restore}
 & A call was made to \tcode{surface::restore} for which no prior call to
 \tcode{surface::save} was made.
 \\
 \tcode{invalid_pop_group}
 & A call was made to \tcode{surface::pop_group} or
 \tcode{surface::pop_group_to_source} for which no prior call to
 \tcode{surface::push_group} was made.
 \\
 \tcode{no_current_point}
 & The operation requires a current point but no current point was set.
 This is usually the result of a call to \tcode{path_builder::rel_curve_to},
 \tcode{path_builder::rel_line_to}, or \tcode{path_builder::rel_move_to} and
 can be corrected by first establishing a current point with a non-"rel"
 member function call such as \tcode{path_builder::move_to}.
 \\
 \tcode{invalid_matrix}
 & A \tcode{matrix_2d} that the operation depends on is invalid. To be valid
 a \tcode{matrix_2d} must be invertible.
 \\
 \tcode{invalid_status}
 & An internal error has occurred.
 \enternote
 This value must only be used when no other \tcode{io2d_error} value is
 appropriate. It signifies that an implementation-specific error
 occurred such as passing a bad native handle as an argument. The
 conditions and circumstances under which this \tcode{io2d_error} value occurs
 are \impldef{\idxcode{status}!\idxcode{null_pointer}}.
 \exitnote
 \\
 \tcode{null_pointer}
 & A null pointer value was unexpectedly encountered.
 \enternote
 The conditions and circumstances under which this \tcode{io2d_error} value 
 occurs 
 are \impldef{\idxcode{status}!\idxcode{null_pointer}}.
 \exitnote
 \\
 \tcode{invalid_string}
 & A UTF-8 string value was expected but the string is not a valid UTF-8
 string.
 \\
 \tcode{invalid_path_data}
 & Invalid data was encountered in a \tcode{path} or a \tcode{path_builder}
 object.
 \enternote
 This status value should only occur when a user directly manipulates the
 internal data of a \tcode{path} or a \tcode{path_builder} object. Creating
 a \tcode{path} object using the \tcode{path_builder} type should never
 result in an invalid path unless the user calls the non-\tcode{const}
 overload of \tcode{path_builder::get_data_ref} and directly manipulates
 the data in an erroneous manner.
 \exitnote
 \\
 \tcode{read_error}
 & An error occurred while attempting to read data from an input stream.
 \\
 \tcode{write_error}
 & An error occurred while attempting to write data to an output stream.
 \\
 \tcode{surface_finished}
 & An attempt was made to use or manipulate a \tcode{surface} object or
 \tcode{surface}-derived object which is no longer valid.
 \enternote
 This can occur due to a previous call to \tcode{surface::finish} or as a
 result of erroneous usage of a native handle.
 \exitnote
 \\
 \tcode{surface_type_mismatch}
 & An operation was attempted on a surface that does not support it.
 \enternote
 Deprecated.
 \exitnote
 \\
 \tcode{pattern_type_mismatch}
 & An operation was attempted on a pattern that does not support it.
 \enternote
 Deprecated.
 \exitnote
  \\
 \tcode{invalid_content}
 & An invalid \tcode{content} value was passed as an argument.
 \enternote
 Deprecated. Implementations should throw an exception of type
 \tcode{std::invalid_argument} instead.
 \exitnote
 \\
 \tcode{invalid_format}
 & An invalid \tcode{format} value was passed as an argument.
 \enternote
 Deprecated. Implementations should throw an exception of type
 \tcode{std::invalid_argument} instead.
 \exitnote
 \\
 \tcode{invalid_visual}
 & Invalid visual.
 \enternote
 Deprecated. Implementations should use \tcode{invalid_status} instead.
 \exitnote
 \\
 \tcode{file_not_found}
 & File not found.
 \\
 \tcode{invalid_dash}
 & An invalid dash value was specified in a call to \tcode{surface::set_dashes}.
 \\
 \tcode{invalid_dsc_comment}
 & Invalid dsc comment.
 \enternote
 Deprecated. Implementations should use \tcode{invalid_status} instead.
 \exitnote
 \\
 \tcode{invalid_index}
 & Invalid index.
 \enternote
 Deprecated. Implementations should throw an exception of type 
 \tcode{std::out_of_range} instead.
 \exitnote
 \\
 \tcode{clip_not_representable}
 & A call was made to \tcode{surface::get_clip_rectangles} when the
 \tcode{surface} object's current clipping region could not be represented
 with rectangles.
 \\
 \tcode{temp_file_error}
 & Temp file error.
 \enternote
 Deprecated. Implementations that require temporary files should use
 \tcode{invalid_status} instead.
 \exitnote
 \\
 \tcode{invalid_stride}
 & An invalid stride value was used. Surface formats may require padding at
 the end of each row of pixel data depending on the implementation and the
 current graphics chipset, if any. Use \tcode{format_stride_for_width} to
 obtain the correct stride value.
 \\
 \tcode{font_type_mismatch}
 & Font type mismatch.
 \enternote
 Deprecated.
 \exitnote
 \\
 \tcode{user_font_immutable}
 & User font immutable.
 \enternote
 Reserved.
 \exitnote
 \\
 \tcode{user_font_error}
 & User font error.
 \enternote
 Reserved.
 \exitnote
 \\
 \tcode{negative_count}
 & Negative count.
 \enternote
 Deprecated.
 \exitnote
 \\
 \tcode{invalid_clusters}
 & A call was made to \tcode{surface::show_text_glyphs} with a
 \tcode{std::vector<text_clusters>} argument that does not properly map
 the UTF-8 \tcode{std::string} code points to the \tcode{std::vector<glyph>}
 glyphs.
 \\
 \tcode{invalid_slant}
 & Invalid slant.
 \enternote
 Deprecated.
 \exitnote
 \\
 \tcode{invalid_weight}
 & Invalid weight.
 \enternote
 Deprecated.
 \exitnote
 \\
 \tcode{invalid_size}
 & The input value for a size parameter is incorrect.
 \enternote
 Typically this is an improper width or height argument. It could be zero or
 it could be a value that exceeds the implementation-defined size limits for
 a surface or a pattern.
 \exitnote
 \\
 \tcode{user_font_not_implemented}
 & User fonts are not implemented.
 \enternote
 Reserved.
 \exitnote
 \\
 \tcode{device_type_mismatch}
 & Device type mismatch.
 \enternote
 Deprecated.
 \exitnote
 \\
 \tcode{device_error}
 & The operation failed. The \tcode{device} encountered an error.
 \enternote
 The conditions and circumstances in which this \tcode{io2d_error} value occurs 
 are 
 \impldef{\idxcode{status}!\idxcode{device_error}}.
 \exitnote
 \\
 \tcode{invalid_mesh_construction}
 & A mesh construction operation on a \tcode{mesh_pattern_builder} object
 failed. Mesh construction operations are only permitted in between a call to
 either \tcode{mesh_pattern_builder::begin_patch} or
 \tcode{mesh_pattern_builder::begin_edit_patch} and
 \tcode{mesh_pattern_builder::end_patch}.
 \\
 \tcode{device_finished}
 & Device finished.
 \enternote
 Deprecated.
 \exitnote
 \\
 \tcode{last_status}
 & This value is unused.
 \enternote
 Deprecated.
 \exitnote
 \\	
\end{libreqtab2}
